Fan variable area nozzles (FVAN) can provide several benefits, such as improving propulsive efficiency, reducing jet noise, increasing fan surge margin, and controlling fan flutter margin, by changing the FVAN area as a function of the current operating condition.  For example, the top of climb flight condition requires a smaller FVAN area whereas take-off requires a larger FVAN area. When analyzing the performance at these steady-state flight conditions, the aforementioned benefits are observed.  However, the gas turbine must operate one steady-state condition to another safely throughout the mission to realize these benefits. 

During a transition from one steady-state condition to another, FVAN must respond in concert with shaft dynamics for performance and safety. If FVAN doesn't respond in accordance with the shaft dynamics, either thrust lapse or stall can occur at different flight conditions. The response time of FVAN depends strongly on actuator dynamics. Lighter actuators with shape memory alloys (SMA) were proposed instead of heavy and complex conventional hydraulic actuators to reduce the fuel burn penalty due to actuator weight when FVAN is on-board. The SMA actuators are more flexible and lighter, but also respond slower than the hydraulic actuators. Therefore, a gas turbine controller must compensate for the slower SMA actuators while controlling either one variable (fuel flow rate with FVAN schedule) or two variables (fuel flow rate and FVAN). This study aims to evaluate how current SMA actuators perform. %If current SMA actuators fail, how slow SMA actuators are acceptable will be determined.  %Do we need this last sentence?

The SMA actuators will be represented by a first order lag with representative time constant values selected from previous work on SMA actuators.  Using a gas turbine model in the small single aisle thrust class, closed-loop controllers will be developed for the cases where FVAN runs with a schedule and when a closed-loop controller directly regulates the FVAN area. Gas turbine transient performance for the two FVAN cases will be analyzed based on thrust lapse and stall margin. A signal with a series of steps, chops, and ramps will be used to test the two developed controllers. If the gas turbine cannot reach 98\% of maximum thrust in less than five seconds (FAA requirement) or keep a safe stall margin throughout the transient operation, the time constant will be reduced to determine the SMA actuator response required to meet the performance and safety requirements.
%A first order lag will represent the SMA actuators and the representative time constant values are selected from the literature on the SMA actuators. Using a gas turbine model in small single aisle thrust class, controllers will be developed for the cases where FVAN runs with a schedule or a controller changes FVAN area directly. Gas turbine transient performance for the two FVAN cases will be analyzed based on thrust lapse and stall margin. A signal with a series of steps, chops, and ramps will be used to test the two developed controllers. If the gas turbine cannot reach 98\% of maximum thrust in less than five seconds (FAA requirement) or keep a safe stall margin throughout the transient operation, the time constant will be reduced to observe how fast the SMA actuator needs to be for performance and safety.